---
title: "free-response experiment"
author: "Alicia Chen"
date: "2025-07-14"
output: html_document
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(tidyboot)
library(glue)
library(lme4)
library(lmerTest)
library(emmeans)
library(simr)
library(GGally)
library(ggimage)
library(reshape2)
library(tidyr)

theme_set(theme_classic(base_size = 15))
options(contrasts = c(unordered = "contr.sum", ordered = "contr.poly"))

# Seed for the bootstrapped CIs, so rendered outputs are reproducible
set.seed(2026)

# Helpers for exporting statistics as LaTeX macros for the manuscript
source(here("analysis/utils/macros.R"))
macros <- list()

# The sensitivity section refits the slowest (maximal) models and is the
# bulk of the render time. Skip it during iteration with
# RUN_SENSITIVITY=false make analysis-...; run the default (true) before
# committing so the sinked outputs and sensitivity.csv stay complete.
run_sensitivity <- tolower(Sys.getenv("RUN_SENSITIVITY", "true")) == "true"
```

# Load data and demographics

```{r}
data <- read_csv(here("data/free-response/selection_trials.csv")) %>%
  mutate(prop_naive = n_naive / (n_blue + n_naive))

demographics <- read_csv(here("data/free-response/exit_survey.csv"))
```

Summary statistics

```{r}
age_summary <- demographics %>%
  summarise(
    age_mean = mean(age, na.rm = TRUE),
    age_sd = sd(age, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE)
  )
age_summary
```



Gender breakdown

```{r}
gender_counts <- demographics %>%
  group_by(gender) %>%
  summarise(n = n())
gender_counts
```

How many participants that understood the instructions?

```{r}
n_participants <- n_distinct(demographics$subject_id)

n_understood <- demographics %>%
  filter(understood == "yes") %>%
  summarise(n_unique = n_distinct(subject_id)) %>%
  pull(n_unique)
n_understood
```

Export demographics as macros

```{r}
n_gender <- function(g) sum(gender_counts$n[gender_counts$gender == g])
macros$expOneN <- fmt_int(n_participants)
macros$expOneNExcluded <- fmt_int(n_participants - n_understood)
macros$expOneNFemale <- fmt_int(n_gender("female"))
macros$expOneNMale <- fmt_int(n_gender("male"))
macros$expOneNNonbinary <- fmt_int(n_gender("nonconforming"))
macros$expOneAgeMin <- fmt_int(age_summary$min_age)
macros$expOneAgeMax <- fmt_int(age_summary$max_age)
macros$expOneAgeMean <- fmt_num(age_summary$age_mean, 1)
macros$expOneAgeSD <- fmt_num(age_summary$age_sd, 1)
```

# Sanity checks

First check the conditions for each subject 
For each subject count the number of unique combinations of (n_blue, n_naive, goal)

```{r}
data %>%
  group_by(subject_id, n_blue, n_naive, goal) %>% 
  summarise(n_trials = n()) 
```


## Learning metrics

Learning metrics: 
- `n_unique`: number of unique tangrams selected per participant (higher = more learning)
- `prop_red`: proportion of times the 'red-specific' tangrams were selected for the 'social' goal (lower = more learning)
- `utt_memory`: mean cosine similarity of the later utterance to the observed later utterance, for trials where there were no naive members in the audience (higher = more learning)

```{r}
learning_metrics <- data %>%
  group_by(subject_id) %>%
  summarise(
    n_unique = n_distinct(selected_tangram),
    prop_red = mean(selected_tangram_group[goal == "social"] == "red_specific", na.rm = TRUE),
    utt_memory = mean(sbert_cosine_later_blue[n_naive == 0], na.rm = TRUE),
  )

learning_metrics
```

Pairplot of learning metrics

```{r}
ggpairs(learning_metrics, columns = c("n_unique", "prop_red", "utt_memory"))
```

We will use the `utt_memory` metric as the main learning metric

Filter out participants who perform in bottom 50% of the distribution of `utt_memory`

```{r}
data_better_performance <- data %>%
  filter(subject_id %in% learning_metrics$subject_id[learning_metrics$utt_memory > quantile(learning_metrics$utt_memory, 0.5)])

write_csv(data_better_performance, here("data/free-response/selection_trials_filtered.csv"))
```


## Tangram selection

How likely are participants to select each of the tangrams, for the 'refer' and 'social' goals? (note that this is averaged across tangram assignments, so doesn't tell us anything about how likely people are to select red-specific tangrams.)

```{r}
tangram_letters <- sort(unique(na.omit(data$selected_tangram)))
tangram_grid <- matrix(tangram_letters, nrow = 3, ncol = 4, byrow = TRUE)

get_counts_matrix <- function(goal) {
  counts <- data %>%
    filter(goal == !!goal) %>%
    count(selected_tangram) %>%
    complete(selected_tangram = tangram_letters, fill = list(n = 0)) %>%
    arrange(selected_tangram)
  matrix(counts$n, nrow = 3, ncol = 4, byrow = TRUE)
}

goals <- c("refer", "social")
plot_data <- lapply(goals, function(goal) {
  mat <- get_counts_matrix(goal)
  df <- melt(mat)
  df$goal <- goal
  df$letter <- as.vector(tangram_grid)
  df
})
plot_data <- do.call(rbind, plot_data)
colnames(plot_data)[1:2] <- c("row", "col")

plot_data$image <- file.path(
  "../../experiments/free-response/stim/tangrams",
  paste0("tangram_", plot_data$letter, ".png")
)

ggplot(plot_data, aes(x = col, y = row, fill = value)) +
  geom_tile(color = "white") +
  geom_image(aes(image = image), size = 0.12, asp = 1) +
  scale_fill_gradient(low = "white", high = "blue") +
  scale_y_reverse() +
  facet_wrap(~goal) +
  theme_minimal(base_size = 15) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(fill = "Count")
```

# Main analyses

## Tangram selection

How often do people select the red-specific tangrams, for the 'refer' and 'social' goals, by proportion naive?

```{r fig.width=10, fig.height=4}
ggplot(data, aes(x = prop_naive, fill = factor(selected_tangram_group, levels = c("blue_specific", "shared", "red_specific")))) +
  geom_histogram(position = "fill", binwidth = 0.1) +
  labs(
    x = "Proportion naive",
    y = "Proportion",
    fill = "Tangram group"
  ) +
  scale_fill_manual(
    values = c("red_specific" = "red", "shared" = "purple", "blue_specific" = "blue"),
    breaks = c("blue_specific", "shared", "red_specific")
  ) +
  facet_wrap(~goal)
```

For the 'social' goal, what are people writing when they choose red-specific tangrams? 

First compare whether the description was more similar to later red, versus blue-specific tangrams to later blue. 

```{r}
red_similarity_summary <- data %>% filter(goal == "social", selected_tangram_group == "red_specific") %>%
  tidyboot_mean(sbert_cosine_later_red, na.rm = TRUE) %>% 
  mutate(selected_tangram_group = "red_specific")

blue_similarity_summary <- data %>% filter(goal == "social", selected_tangram_group == "blue_specific") %>%
  tidyboot_mean(sbert_cosine_later_blue, na.rm = TRUE) %>% 
  mutate(selected_tangram_group = "blue_specific")

similarity_summary <- rbind(red_similarity_summary, blue_similarity_summary)
```

```{r}
ggplot(similarity_summary, aes(x = selected_tangram_group, y = empirical_stat, color = selected_tangram_group)) +
  geom_jitter(
    data = data %>% filter(goal == "social", selected_tangram_group %in% c("red_specific", "blue_specific")),
    aes(
      x = selected_tangram_group,
      y = ifelse(selected_tangram_group == "red_specific", sbert_cosine_later_red, sbert_cosine_later_blue),
      color = selected_tangram_group
    ),
    width = 0.15, alpha = 0.05, size = 2, inherit.aes = FALSE
  ) +
  geom_point(position = position_dodge(width = 0.03), size = 2) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 2, width = 0) +
  scale_color_manual(values = c("red_specific" = "red", "blue_specific" = "blue")) +
  labs(x = "Tangram group", y = "Cosine similarity to later", title = "Social goal", subtitle = "is blue more similar to later blue, or red more similar to later red?") 
```

The above plot shows that when participants choose blue-specific tangrams, they write descriptions that are similar to the blue group's converged utterances for those tangrams — but when participants select red-specific tangrams, they do NOT write descriptions that are similar to the red group's converged utterances for those tangrams

Next, what exactly are they writing? When people choose red-specific tangrams, are they writing something that's similar to the blue-group expressions? (i.e. do they talk about things that are not the tangram they selected, and instead provide descriptions that are identifiable by the blue group?)

For each time a red-specific tangram is selected for the social goal: 
- calculate similarity of the written description to each of the blue-group converged expressions for all of the blue-group tangrams (blue-specific + shared)
- calculate similarity of the written description to each of the red-group converged expressions for all of the red-group tangrams (red-specific + shared)
- select the most similar later description among each of the blue-group tangrams
- select the most similar later description among each of the red-group tangrams

We need to do this in a python script so we can use the SBERT model, but first find the converged expressions for each tangram

```{r}
converged_expressions <- data %>%
  group_by(item_id, counterbalance, selected_tangram, selected_tangram_group, selected_tangram_later_red, selected_tangram_later_blue) %>%
  summarise()

# save to csv
write_csv(converged_expressions, here("data/free-response/converged_expressions.csv"))
```

This is where we run `red_social_similarity.py` and the output of that is in `red_social_similarity_results.csv`

```{r}
red_social_similarity <- read_csv(here("analysis/free-response/red_social_similarity_results.csv"))
```

```{r}
ggplot(red_social_similarity, aes(x = max_blue_sim, y = max_red_sim)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") + 
  lims(x = c(0, 1), y = c(0, 1)) +
  labs(x = "Max similarity to blue-group converged", y = "Max similarity to red-group converged", title = "Social goal", subtitle = "Descriptions for selected red-specific (vs blue-specific)") + 
  facet_wrap(~selected_tangram_group) +
  coord_fixed() 
```

```{r}
ggplot(red_social_similarity, aes(x = mean_blue_sim, y = mean_red_sim)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") + 
  labs(x = "Mean similarity to blue-group converged", y = "Mean similarity to red-group converged", title = "Social goal", subtitle = "Descriptions for selected red-specific (vs blue-specific)") + 
  facet_wrap(~selected_tangram_group) +
  coord_fixed() 
```


```{r}
summary_max_blue <- red_social_similarity %>%
  group_by(selected_tangram_group) %>%
  tidyboot_mean(max_blue_sim, na.rm = TRUE) %>%
  mutate(similarity_to = "blue-group")

summary_max_red <- red_social_similarity %>%
  group_by(selected_tangram_group) %>%
  tidyboot_mean(max_red_sim, na.rm = TRUE) %>%
  mutate(similarity_to = "red-group")

summary_max <- rbind(summary_max_blue, summary_max_red)
```

```{r}
ggplot(summary_max, aes(x = similarity_to, y = empirical_stat, color = similarity_to)) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_manual(values = c("blue-group" = "blue", "red-group" = "red")) +
  labs(x = "similarity_to", y = "Cosine similarity", title = "Social goal", subtitle = "Descriptions for selected red-specific (vs blue-specific)") + 
  facet_wrap(~selected_tangram_group) 
```

The above plot shows that in general, people write blue group labels when they select blue-specific tangrams, and they write red-group labels when they select red-specific tangrams. BUT, while participants are extremely likely to write blue-group labels for blue-specific tangrams (recall that the social goal is only directed to the blue group), they are comparatively less likely to write red-group labels for red-specific tangrams.

Remake plot with individual points

```{r}
ggplot(summary_max, aes(x = similarity_to, y = empirical_stat, color = similarity_to)) +
  geom_jitter(
    data = red_social_similarity %>% filter(selected_tangram_group %in% c("red_specific", "blue_specific")) %>% pivot_longer(cols = c(max_blue_sim, max_red_sim), names_to = "similarity_to", values_to = "value") %>% mutate(similarity_to = ifelse(similarity_to == "max_blue_sim", "blue-group", "red-group")),
    aes(
      x = similarity_to,
      y = value,
      color = similarity_to
    ),
    width = 0.2, alpha = 0.03, size = 2, inherit.aes = FALSE
  ) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_manual(values = c("blue-group" = "blue", "red-group" = "red")) +
  labs(x = "similarity_to", y = "Cosine similarity", title = "Social goal", subtitle = "Descriptions for selected red-specific (vs blue-specific)") + 
  facet_wrap(~selected_tangram_group) 
```



##  Utterance length

Plot utterance length by proportion naive and goal

```{r}
utt_length_summary <- data %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(utt_length, na.rm = TRUE)
```

```{r}
ggplot(utt_length_summary, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "Utterance length", color = "Goal") 
```

Add individual points

```{r fig.width=10}

ggplot(utt_length_summary, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_jitter(
    data = data,
    aes(x = prop_naive, y = utt_length, color = goal),
   alpha = 0.1, size = 2, inherit.aes = FALSE,
   position = position_jitterdodge(dodge.width = 0.05)
  ) +
  geom_point(position = position_dodge(width = 0.05), size = 3) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.05), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "Utterance length", color = "Goal") 
```

Log transform utterance lengths 

```{r}
utt_length_summary_log <- data %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(log(utt_length), na.rm = TRUE)
```

```{r}
ggplot(utt_length_summary_log, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "Log utterance length", color = "Goal")
```

```{r fig.width=10}
ggplot(utt_length_summary_log, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_jitter(
    data = data,
    aes(x = prop_naive, y = log(utt_length), color = goal),
   alpha = 0.1, size = 2, inherit.aes = FALSE,
   position = position_jitterdodge(dodge.width = 0.05, jitter.height = 0.1)
  ) +
  geom_point(position = position_dodge(width = 0.05), size = 3) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.05), size = 1.7, width = 0) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "Log utterance length", color = "Goal") 
```



## Cosine similarity to later utterance

Plot cosine similarity to later utterance by proportion naive and goal

```{r}
cos_diff_summary <- data %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(1 - sbert_cosine_later_blue, na.rm = TRUE)
```

```{r}
ggplot(cos_diff_summary, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "1 - cosine similarity to later", color = "Goal") 
```

Add individual points

```{r fig.width=10}
ggplot(cos_diff_summary, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_jitter(
    data = data,
    aes(x = prop_naive, y = 1 - sbert_cosine_later_blue, color = goal),
   alpha = 0.1, size = 2, inherit.aes = FALSE,
   position = position_jitterdodge(dodge.width = 0.05, jitter.height = 0.01)
  ) +
  geom_point(position = position_dodge(width = 0.05), size = 3) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.05), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "1 - cosine similarity to later", color = "Goal") 
```



## Filter out trials where participants selected red-specific tangrams

```{r}
data_no_red <- data %>%
  filter(selected_tangram_group != "red_specific")
```

Remake utt length and semantic similarity without red-specific tangrams

```{r}
utt_length_summary_no_red <- data_no_red %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(utt_length, na.rm = TRUE)
```

```{r}
cos_diff_summary_no_red <- data_no_red %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(1 - sbert_cosine_later_blue, na.rm = TRUE)
```

Plots

```{r}
ggplot(utt_length_summary_no_red, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "Utterance length", title = "Without red-specific choices", color = "Goal") 
```

```{r}
ggplot(cos_diff_summary_no_red, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "1 - cosine similarity to later", title = "Without red-specific choices", color = "Goal") 
```

# Data analysis

```{r}
data <- data %>%
  mutate(
    tangram_choice = ifelse(selected_tangram_group == "red_specific", 1, 0),
    goal = factor(goal, levels = c("social", "refer")),
    prop_naive = as.numeric(prop_naive)
  )

# Helpers for extracting the manuscript-quoted statistics from model summaries.
# glmer coefficients are Wald z tests; lmer coefficients (via lmerTest) are
# Satterthwaite t tests, so those also export the (data-dependent) df.
glmer_coef_macros <- function(mod, term) {
  co <- summary(mod)$coefficients[term, ]
  list(B = fmt_num(co[["Estimate"]]), Z = fmt_num(co[["z value"]]), P = fmt_p(co[["Pr(>|z|)"]]))
}
lmer_coef_macros <- function(mod, term) {
  co <- summary(mod)$coefficients[term, ]
  list(
    B = fmt_num(co[["Estimate"]]), T = fmt_num(co[["t value"]]),
    Df = fmt_num(co[["df"]], 1), P = fmt_p(co[["Pr(>|t|)"]])
  )
}
# Extract one goal x degree row of an emtrends summary as B/Z/P macro values
# (with n > 3000 observations, emmeans uses the asymptotic z scale for lmer too)
trend_macros <- function(trends, deg, g) {
  row <- as.data.frame(trends)
  row <- row[row$degree == deg & row$goal == g, ]
  list(B = fmt_num(row$prop_naive.trend), Z = fmt_num(row$z.ratio), P = fmt_p(row$p.value))
}
```

Tangram choice model, with a linear term for proportion naive (kept for the likelihood ratio test against the polynomial model below)

```{r}
# This sink intentionally opens without append = TRUE: it is the first sink in
# the document, so each full render rebuilds model_outputs.txt from scratch.
sink(here("analysis/free-response/model_outputs.txt"))

# The random-effects structure follows the preregistered contingency plan
# (drop the random interactions, then the slopes, and/or run the Bayesian
# equivalent). The maximal specification, (prop_naive + goal | subject_id) +
# (prop_naive + goal | item_id) + (prop_naive + goal | selected_tangram),
# produced a singular fit, and the boundary components were removed one at a
# time: the tangram-set (item_id, 3 levels) slope correlations were degenerate
# (+/-1) and its intercept variance was then estimated at zero, so item_id was
# removed; the by-participant goal slope was perfectly correlated (-1) with
# the prop_naive slope, so it was removed; and the by-tangram prop_naive
# slope was then perfectly correlated (+1) with the by-tangram intercept, so
# it was removed. The non-singular structure below retains the by-participant
# prop_naive slopes and the by-tangram goal slopes. The goal estimates are
# essentially unchanged under alternative specifications (see the sensitivity
# analysis).
mod <- glmer(
  tangram_choice ~ prop_naive * goal
  + (prop_naive | subject_id)
  + (goal | selected_tangram),
  data = data,
  family = 'binomial',
  control = glmerControl(optimizer = "bobyqa")
)

summary(mod)

sink()
```

try with polynomial term

```{r}
sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\n")

# Same pruned random-effects structure as the linear model above (see the
# comment there), so the two models are nested for the likelihood ratio test.
mod_poly <- glmer(
  tangram_choice ~ poly(prop_naive, 2) * goal
  + (prop_naive | subject_id)
  + (goal | selected_tangram),
  data = data,
  family = 'binomial',
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)
summary(mod_poly)

sink()

macros <- macros %>%
  add_macros("expOneTangramSocial", glmer_coef_macros(mod_poly, "goal1"))
```
Likelihood ratio test

```{r}
sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\n")

print(anova(mod, mod_poly, test = "Chisq"))

sink()
```
The model with the polynomial fit is a better fit to the data

```{r}
sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\n")

trends_tangram <- emtrends(mod_poly, ~ goal, var = "prop_naive", type = "response", max.degree = 2) %>%
  summary(infer = T)
print(trends_tangram)

sink()

macros <- macros %>%
  add_macros("expOneTangramLin", trend_macros(trends_tangram, "linear", "refer")) %>%
  add_macros("expOneTangramQuad", trend_macros(trends_tangram, "quadratic", "refer"))
```


```{r}
sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\n")

# The random-effects structure follows the preregistered contingency plan
# (drop the random interactions, then the slopes, and/or run the Bayesian
# equivalent). The maximal specification, with (prop_naive + goal | item_id),
# produced a singular fit: the tangram-set (item_id, 3 levels) slope
# correlations were all degenerate (+1), and with the slopes removed the
# item_id intercept variance was estimated at exactly zero, so item_id was
# removed. The non-singular structure below retains the full by-participant
# and by-tangram slopes. The goal estimates are essentially unchanged under
# alternative specifications (see the sensitivity analysis).
mod_utt_length <- lmer(
  utt_length ~ poly(prop_naive, 2) * goal
  + (prop_naive + goal | subject_id)
  + (prop_naive + goal | selected_tangram),
  control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
  data = data
)

summary(mod_utt_length)

sink()

macros <- macros %>%
  add_macros("expOneLengthSocial", lmer_coef_macros(mod_utt_length, "goal1"))
macros$expOneLengthInterP <- fmt_p(
  summary(mod_utt_length)$coefficients["poly(prop_naive, 2)1:goal1", "Pr(>|t|)"]
)
```

```{r}
sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\n")

trends_length <- emtrends(mod_utt_length, ~ goal, var = "prop_naive", type = "response", max.degree = 2) %>%
  summary(infer = T)
print(trends_length)

sink()

macros <- macros %>%
  add_macros("expOneLengthLin", trend_macros(trends_length, "linear", "refer")) %>%
  add_macros("expOneLengthQuad", trend_macros(trends_length, "quadratic", "refer"))
```

```{r}
sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\n")

# The random-effects structure follows the preregistered contingency plan
# (drop the random interactions, then the slopes, and/or run the Bayesian
# equivalent). The previous specification, with (goal | item_id), produced a
# singular fit: the tangram-set (item_id, 3 levels) variance components were
# estimated at essentially zero with a degenerate (-1) correlation, so item_id
# was removed. The structure below retains the full by-participant and
# by-tangram components. The goal estimates are essentially unchanged under
# alternative specifications (see the sensitivity analysis).
mod_sbert <- lmer(
  1 - sbert_cosine_later_blue ~ poly(prop_naive, 2) * goal
  + (prop_naive + goal | subject_id)
  + (goal | selected_tangram),
    control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e6)),
  data = data
)

summary(mod_sbert)

sink()

macros <- macros %>%
  add_macros("expOneSemSocial", lmer_coef_macros(mod_sbert, "goal1"))
macros$expOneSemInterP <- fmt_p(
  summary(mod_sbert)$coefficients["poly(prop_naive, 2)1:goal1", "Pr(>|t|)"]
)
```


```{r}
sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\n")

trends_sbert <- emtrends(mod_sbert, ~ goal, var = "prop_naive", type = "response", max.degree = 2) %>%
  summary(infer = T)
print(trends_sbert)

sink()

macros <- macros %>%
  add_macros("expOneSemLin", trend_macros(trends_sbert, "linear", "refer")) %>%
  add_macros("expOneSemQuad", trend_macros(trends_sbert, "quadratic", "refer"))
```

## Sensitivity analysis

The social-goal main effect for each of the three models under alternative random-effects specifications, from intercepts-only up to the richest structure previously fit. The richest specifications are singular (boundary fits) and are shown only to demonstrate that the fixed-effects conclusions do not depend on the choice of structure.

```{r, eval = run_sensitivity}
sensitivity_specs <- list(
  tangram_choice = list(
    response = "tangram_choice",
    "intercepts only" = ~ (1 | subject_id) + (1 | selected_tangram),
    "final (reported)" = ~ (prop_naive | subject_id) + (goal | selected_tangram),
    "maximal (singular)" = ~ (prop_naive + goal | subject_id) + (prop_naive + goal | item_id) + (prop_naive + goal | selected_tangram)
  ),
  utt_length = list(
    response = "utt_length",
    "intercepts only" = ~ (1 | subject_id) + (1 | selected_tangram),
    "final (reported)" = ~ (prop_naive + goal | subject_id) + (prop_naive + goal | selected_tangram),
    "maximal (singular)" = ~ (prop_naive + goal | subject_id) + (prop_naive + goal | item_id) + (prop_naive + goal | selected_tangram)
  ),
  semantic_distance = list(
    response = "1 - sbert_cosine_later_blue",
    "intercepts only" = ~ (1 | subject_id) + (1 | selected_tangram),
    "final (reported)" = ~ (prop_naive + goal | subject_id) + (goal | selected_tangram),
    "previous (singular)" = ~ (prop_naive + goal | subject_id) + (goal | item_id) + (goal | selected_tangram)
  )
)

fit_sensitivity <- function(model_name, spec_name) {
  spec <- sensitivity_specs[[model_name]]
  re_terms <- paste(deparse(spec[[spec_name]][[2]]), collapse = " ")
  f <- as.formula(glue("{spec$response} ~ poly(prop_naive, 2) * goal + {re_terms}"))
  if (model_name == "tangram_choice") {
    m <- glmer(f, data = data, family = "binomial",
               control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
  } else {
    m <- lmer(f, data = data,
              control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e6)))
  }
  co <- summary(m)$coefficients["goal1", ]
  tibble(
    model = model_name,
    specification = spec_name,
    singular = isSingular(m),
    estimate = co[["Estimate"]],
    SE = co[["Std. Error"]],
    stat = ifelse(model_name == "tangram_choice", co[["z value"]], co[["t value"]]),
    p = ifelse(model_name == "tangram_choice", co[["Pr(>|z|)"]], co[["Pr(>|t|)"]])
  )
}

sensitivity <- map_dfr(names(sensitivity_specs), function(model_name) {
  spec_names <- setdiff(names(sensitivity_specs[[model_name]]), "response")
  map_dfr(spec_names, function(spec_name) fit_sensitivity(model_name, spec_name))
})

write_csv(sensitivity, here("analysis/free-response/sensitivity.csv"))

sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\nSENSITIVITY ANALYSIS: social-goal main effect (goal1 coefficient) under\n")
cat("alternative random-effects specifications (see sensitivity.csv):\n\n")
print(as.data.frame(sensitivity))
sink()
```

# Remake plots with better performing participants

Recalculate summary statistics with better performing participants

```{r}
utt_length_summary_better_performance <- data_better_performance %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(utt_length, na.rm = TRUE)
```

```{r}
cos_diff_summary_better_performance <- data_better_performance %>%
  filter(!is.na(sbert_cosine_later_blue)) %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(1 - sbert_cosine_later_blue, na.rm = TRUE)
```

Plots

```{r fig.width=10, fig.height=4}
ggplot(data_better_performance, aes(x = prop_naive, fill = factor(selected_tangram_group, levels = c("blue_specific", "shared", "red_specific")))) +
  geom_histogram(position = "fill", binwidth = 0.1) +
  labs(
    x = "Proportion naive",
    y = "Proportion",
    title = "Better memory participants",
    fill = "Tangram group"
  ) +
  scale_fill_manual(
    values = c("red_specific" = "red", "shared" = "purple", "blue_specific" = "blue"),
    breaks = c("blue_specific", "shared", "red_specific")
  ) +
  facet_wrap(~goal)
```

```{r}
ggplot(utt_length_summary_better_performance, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "Utterance length", title = "Better memory participants", color = "Goal") 
```

```{r}
ggplot(cos_diff_summary_better_performance, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_point(position = position_dodge(width = 0.03), size = 3) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.03), size = 1.7, width = 0) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Proportion naive", y = "1 - cosine similarity to later", title = "Better memory participants", color = "Goal") 
```

## Other tests

Correlate how well people learned the descriptions, with size of 'goal' effect for each participant

```{r}
eff_cor_tangram <- data %>%
  group_by(subject_id) %>%
  summarise(
    goal_effect = mean(tangram_choice[goal == "refer"], na.rm = TRUE) - mean(tangram_choice[goal == "social"], na.rm = TRUE),
    utt_memory = mean(sbert_cosine_later_blue[n_naive == 0], na.rm = TRUE)
  ) 

eff_cor_sbert <- data %>%
  group_by(subject_id) %>%
  summarise(
    goal_effect = mean(sbert_cosine_later_blue[goal == "social"], na.rm = TRUE) - mean(sbert_cosine_later_blue[goal == "refer"], na.rm = TRUE),
    utt_memory = mean(sbert_cosine_later_blue[n_naive == 0], na.rm = TRUE)
  ) 

eff_cor_length <- data %>%
  group_by(subject_id) %>%
  summarise(
    goal_effect = mean(utt_length[goal == "refer"], na.rm = TRUE) - mean(utt_length[goal == "social"], na.rm = TRUE),
    goal_effect_log = mean(log(utt_length[goal == "refer"]), na.rm = TRUE) - mean(log(utt_length[goal == "social"]), na.rm = TRUE),
    utt_memory = mean(sbert_cosine_later_blue[n_naive == 0], na.rm = TRUE)
  ) 
```

Calculate correlations

```{r}
sink(here("analysis/free-response/model_outputs.txt"), append = TRUE)
cat("\n\n")

cor_tangram <- cor.test(eff_cor_tangram$utt_memory, eff_cor_tangram$goal_effect)
print(cor_tangram)

cat("\n")

cor_length <- cor.test(eff_cor_length$utt_memory, eff_cor_length$goal_effect)
print(cor_length)

cat("\n")

cor_sbert <- cor.test(eff_cor_sbert$utt_memory, eff_cor_sbert$goal_effect)
print(cor_sbert)

sink()

cor_macros <- function(ct) {
  list(
    R = fmt_num(ct$estimate),
    T = fmt_num(ct$statistic),
    Df = fmt_int(ct$parameter),
    P = fmt_p(ct$p.value)
  )
}
macros <- macros %>%
  add_macros("expOneCorrTangram", cor_macros(cor_tangram)) %>%
  add_macros("expOneCorrLength", cor_macros(cor_length)) %>%
  add_macros("expOneCorrSem", cor_macros(cor_sbert))
```





Plot

```{r}
ggplot(eff_cor_sbert, aes(x = utt_memory, y = goal_effect)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "Memory of later blue utterances", y = "Goal effect - semantic similarity")
```

```{r}
ggplot(eff_cor_length, aes(x = utt_memory, y = goal_effect)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "Memory of later blue utterances", y = "Goal effect - utterance length")
```

```{r}
ggplot(eff_cor_length, aes(x = utt_memory, y = goal_effect_log)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "Memory of later blue utterances", y = "Goal effect - log utterance length")
```

```{r}
ggplot(eff_cor_tangram, aes(x = utt_memory, y = goal_effect)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "Memory of later blue utterances", y = "Goal effect - tangram choice")
```

# Export statistics as LaTeX macros

Write out every statistic quoted in the manuscript as a LaTeX macro, so the paper never hardcodes a computed value (`make manuscript-stats` copies these into the manuscript).

```{r}
write_macros(macros, here("analysis/free-response/macros.tex"))

# Export the final random-effects structure of each fitted model, straight
# from the fitted objects, for the manuscript's appendix table.
re_structures <- tibble::tibble(
  model = c("Tangram choice", "Utterance length", "Semantic distance"),
  random_effects = vapply(list(mod_poly, mod_utt_length, mod_sbert), re_terms, character(1))
)
write_csv(re_structures, here("analysis/free-response/re_structures.csv"))
```
















