---
title: "earlier-later 3rd party experiment data analysis"
author: "Alicia Chen"
date: "2024-12-10"
output:
  html_document: default
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(tidyboot)
library(glue)
library(lme4)
library(lmerTest)
library(emmeans)
library(simr)

theme_set(theme_classic(base_size = 20))
options(contrasts = c(unordered = "contr.sum", ordered = "contr.poly"))

# Seed for the bootstrapped CIs, so rendered outputs are reproducible
set.seed(2026)

# Helpers for exporting statistics as LaTeX macros for the manuscript
source(here("analysis/utils/macros.R"))
macros <- list()

# The sensitivity section refits the slowest (maximal) models and is the
# bulk of the render time. Skip it during iteration with
# RUN_SENSITIVITY=false make analysis-...; run the default (true) before
# committing so the sinked outputs and sensitivity.csv stay complete.
run_sensitivity <- tolower(Sys.getenv("RUN_SENSITIVITY", "true")) == "true"
```

# Data, participants, and sanity checks

Load data

```{r}
selection_trials <- read_csv(here("data/earlier-later/selection_trials.csv")) %>%
  group_by(subject_id) %>%
  mutate(trial_num = row_number()) %>%
  rename(response.unique = response.shared) %>%
  select(subject_id, item_id, trial_num, everything()) %>%
  ungroup()

selection_trials$condition <- with(selection_trials, interaction(audience, goal, sep = " "))
selection_trials$tangram <- selection_trials$option1.tangram
selection_trials$shared <- selection_trials$option1.shared

exit_survey <-read_csv(here("data/earlier-later/exit_survey.csv"))
```

How many participants?

```{r}
n_participants <- n_distinct(selection_trials$subject_id)

n_understood <- exit_survey %>%
  filter(understood == "yes") %>%
  nrow()

print(glue("we have {n_participants} total participants, of which {n_understood} said they understood the instructions"))

# Exclude the participants who did not understand the instructions
selection_trials <- selection_trials %>%
  filter(understood == "yes")
```

Demographics

```{r}
gender_counts <- exit_survey %>%
  group_by(gender) %>%
  count()
gender_counts

age_summary <- exit_survey %>%
  summarize(mean_age = mean(age), sd_age = sd(age), min_age = min(age),
    max_age = max(age))
age_summary

n_gender <- function(g) sum(gender_counts$n[gender_counts$gender == g])
macros$expThreeN <- fmt_int(n_participants)
macros$expThreeNExcluded <- fmt_int(n_participants - n_understood)
macros$expThreeNFemale <- fmt_int(n_gender("female"))
macros$expThreeNMale <- fmt_int(n_gender("male"))
macros$expThreeNNonbinary <- fmt_int(n_gender("nonconforming"))
macros$expThreeAgeMin <- fmt_int(age_summary$min_age)
macros$expThreeAgeMax <- fmt_int(age_summary$max_age)
macros$expThreeAgeMean <- fmt_num(age_summary$mean_age, 1)
macros$expThreeAgeSD <- fmt_num(age_summary$sd_age, 1)
```

## Counterbalancing assignments

How many participants were in each item_id, counterbalancing assignment?

```{r}
selection_trials %>%
  group_by(item_id, counterbalance) %>%
  summarize(n = n_distinct(subject_id))
```

## How many trials per condition?

There should be 12 trials per audience-goal condition

```{r}
selection_trials %>%
  filter(type == "main") %>%
  group_by(subject_id, audience, goal) %>%
  count()
```

There should be 9 'refer' goal baseline trials, and 6 'social' goal baseline trials

```{r}
selection_trials %>%
  filter(type == "baseline") %>%
  group_by(subject_id, audience, goal) %>%
  count()
```

# Main trials

Check how many observations per 2AFC pair (for model fitting)

```{r}
observation.count <- selection_trials %>%
    filter(type == "main") %>%
    group_by(item_id, counterbalance, tangram, shared, audience_group, option1.label, option2.label) %>%
    rename(earlier_label = option1.label, later_label = option2.label) %>%
    count()

# to csv
write_csv(observation.count, here("data/earlier-later/labels.csv"))
```

Clean selection trials and output for model fitting

```{r}
selection_trials_clean <- selection_trials %>%
  filter(type == "main") %>%
  select(subject_id, item_id, counterbalance, condition, goal, audience, audience_group, tangram, shared, response.earlier) 

write_csv(selection_trials_clean, here("data/earlier-later/selection_trials_clean.csv"))
```


Calculate means

```{r}
main.trials <- selection_trials %>%
  filter(type == "main") %>%
  mutate(response.earlier = ifelse(response.earlier == "earlier", 1, 0))

main.trials.means <- main.trials %>%
  group_by(condition, shared) %>%
  tidyboot_mean(response.earlier, na.rm = T)

# Per-participant means for the spaghetti lines; plain means (no bootstrap)
# because only the point estimates are used downstream.
main.trials.means.subj <-
  main.trials %>%
  group_by(subject_id, condition, shared) %>%
  summarize(empirical_stat = mean(response.earlier, na.rm = T), .groups = "drop")
```

Plot

```{r fig.width=8, fig.height=4}
ggplot(main.trials,
       aes(x = condition, y = response.earlier, color = shared)) +
  geom_point(
    data = main.trials.means,
    position = position_dodge(width = 0.3),
    aes(x = condition, y = empirical_stat, color = shared),
    size = 3,
    fill = "black",
    inherit.aes = FALSE
  ) +
  geom_errorbar(
    data = main.trials.means,
    position = position_dodge(width = 0.3),
    aes(
      x = condition,
      ymin = ci_lower,
      ymax = ci_upper,
      color = shared
    ),
    size = 1.5,
    width = 0.05,
    inherit.aes = FALSE
  ) +
  geom_line(
    data = main.trials.means.subj,
    aes(x = condition, y = empirical_stat, group = interaction(subject_id, shared), color = shared),
    position = position_dodge(width = 0.3),
    inherit.aes = FALSE, 
    alpha = 0.15
  ) +
  geom_hline(yintercept = 0.5,
             color = "red",
             linetype = "dashed") +
  labs(x = "condition", y = "earlier or later label") +
  scale_y_continuous(breaks = c(0, 1),
                     labels = c("later", "earlier")) +
  scale_color_discrete(name = "label", labels = c("shared", "group-specific"))
```



# Visualize things about baseline trials

## "Refer" goal

For the "refer" goal, people should choose the seen label over the unseen label, both when talking to one group and when talking to either group

```{r}
baseline.trials.refer <- selection_trials %>%
  filter(type == "baseline", goal == "refer") %>%
  mutate(seen = case_when(
    response.group == "unseen" ~ 0,
    response.unique == "shared" ~ 1
  ))

baseline.trials.refer.means <- baseline.trials.refer %>%
  group_by(option1.label, option2.label) %>%
  tidyboot_mean(seen, na.rm = TRUE) %>%
  mutate(`seen - unseen` = paste(option1.label, option2.label, sep = " - ")) %>%
  arrange(empirical_stat)

# Relevel AFTER arrange
baseline.trials.refer.means$`seen - unseen` <- factor(
  baseline.trials.refer.means$`seen - unseen`,
  levels = baseline.trials.refer.means$`seen - unseen`
)
```


Plot

```{r}
ggplot(baseline.trials.refer.means, aes(x = `seen - unseen`, y = empirical_stat)) +
  geom_point(size = 3, fill = "black") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), size = 1.5, width = 0.2) +
  geom_hline(yintercept = 0.5, color = "red", linetype = "dashed") +
  labs(x = "seen label - unseen label", y = "P(seen label)", title = "Study 1b manipulation checks, refer goal") +
  scale_y_continuous(limits = c(0, 1)) +
  theme(text = element_text(size = 15), axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
baseline.refer.means <- baseline.trials.refer %>%
  tidyboot_mean(seen, na.rm = T)

baseline.refer.means
```

## "Social" goal

For the "social" goal, people should choose the tangram belonging to the audience group

```{r}
baseline.trials.social <- selection_trials %>%
  filter(type == "baseline", goal == "social") %>%
  mutate(group_label = ifelse(audience_group == response.group, 1, 0))

baseline.trials.social.means <- baseline.trials.social %>%
  group_by(option1.label, option2.label) %>% 
  tidyboot_mean(group_label, na.rm = T) %>%
  mutate(goal = "social", 
         `red - blue` = paste(option1.label, option2.label, sep = " - ")) %>%
  arrange(empirical_stat)

# Relevel AFTER arrange
baseline.trials.social.means$`red - blue` <- factor(
  baseline.trials.social.means$`red - blue`,
  levels = baseline.trials.social.means$`red - blue`
)
```

Plot

```{r}
ggplot(baseline.trials.social.means, aes(x = `red - blue`, y = empirical_stat)) +
  geom_point(size = 3, fill = "black") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), size = 1.5, width = 0.2) +
  geom_hline(yintercept = 0.5, color = "red", linetype = "dashed") +
  labs(x = "red label - blue label", y = "P(correct group)", title = "Manipulation checks, social goal") +
  scale_y_continuous(limits = c(0, 1)) +
  theme(text = element_text(size = 15), axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
baseline.social.means <- baseline.trials.social %>%
  tidyboot_mean(group_label, na.rm = T) %>%
  mutate(goal = "social")

baseline.social.means
```
### Correlate social baseline performance with effect size of social goal

```{r}
pct.correct.social <- baseline.trials.social %>%
  group_by(subject_id) %>%
  summarize(pct_correct = mean(group_label, na.rm = TRUE))

# effect size is mean response.earlier for "one refer" minus mean response.earlier for "one social"
# calculate effect size for each participant
one_refer_means <- main.trials %>%
  filter(condition == "one refer") %>%
  group_by(subject_id) %>%
  summarize(one_refer_mean = mean(response.earlier, na.rm = TRUE))

one_social_means <- main.trials %>%
  filter(condition == "one social") %>%
  group_by(subject_id) %>%
  summarize(one_social_mean = mean(response.earlier, na.rm = TRUE))

# merge the two dataframes
merged_data <- one_refer_means %>%
  inner_join(one_social_means, by = "subject_id")

# calculate effect size
merged_data <- merged_data %>%
  mutate(effect_size = one_refer_mean - one_social_mean)

# merge with pct_correct
merged_data <- merged_data %>%
  inner_join(pct.correct.social, by = "subject_id")

# correlate effect size with pct_correct

# This sink intentionally opens without append = TRUE: it is the first sink in
# the document, so each full render rebuilds model_outputs.txt from scratch.
sink(here("analysis/earlier-later/model_outputs.txt"))

cor.test(merged_data$effect_size, merged_data$pct_correct)

sink()
```
plot scatterplot of effect size vs pct_correct
```{r}
ggplot(merged_data, aes(x = effect_size, y = pct_correct)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Effect size (one refer - one social)", y = "Pct correct (social baseline)", title = "Effect size vs. social baseline performance") 
```


## Put the baseline trials together

Make a bar plot where x axis is participant, and for each participant it shows total baseline trials correct, colored by \# trials selected "seen" for the refer goal, and "group label" for the social goal

```{r}
baseline.trials.refer.total <- baseline.trials.refer %>%
  group_by(subject_id) %>%
  summarize(total = sum(seen))

baseline.trials.social.total <- baseline.trials.social %>%
  group_by(subject_id) %>%
  summarize(total = sum(group_label))

baseline.trials.total <- baseline.trials.refer.total %>%
  inner_join(baseline.trials.social.total, by = "subject_id") %>%
  mutate(refer = total.x, social = total.y) %>%
  select(subject_id, refer, social) %>%
  mutate(total_correct = refer + social) %>%
  arrange(-total_correct, -social) %>%  # First order by total correct, then by social
  mutate(subject_id = factor(subject_id, levels = unique(subject_id))) %>%  # Preserve the ordering
  pivot_longer(cols = c(refer, social), names_to = "goal", values_to = "total")
```

Make bar plot

```{r}
ggplot(baseline.trials.total, aes(x = subject_id, y = total, fill = goal)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "participant", y = "total correct", fill = "trial type", title = "baseline trials") +
  theme(axis.text.x = element_blank()) + 
  scale_fill_brewer(palette = "Set2") 
```

# Data analysis

## main trials

```{r}
sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\n")

# The random-effects structure follows the preregistered simplification ladder
# ("drop the random interactions, then the slopes, and/or run the Bayesian
# equivalent"). The preregistered maximal specification,
# (condition * shared | subject_id) + (condition | item_id/tangram), produced
# a singular fit: the tangram-set (item_id) variance components were estimated
# at ~zero with degenerate (+/-1) correlations (only 3 tangram sets cannot
# support them), the by-tangram condition slopes were near-degenerate, and the
# by-subject condition slopes were almost perfectly correlated (r = -0.99),
# i.e. rank-deficient. Following the ladder, we removed the item_id components
# and the by-tangram condition slopes, then dropped the random condition:shared
# interaction; the by-subject condition slopes remained degenerate, so we
# decomposed them: condition is the audience x goal interaction, and its
# by-subject variance was concentrated along the audience (either vs. known
# group) contrast, with the by-subject goal slope variance estimated at ~zero.
# Removing the zero/degenerate components yields the non-singular structure
# below, which retains by-subject slopes for audience and tangram type and
# by-tangram intercepts. Estimates under alternative specifications are
# essentially identical (see the sensitivity analysis).
mod <- glmer(response.earlier ~ condition * shared + (audience + shared | subject_id) + (1 | item_id:tangram), data = main.trials, family = "binomial", control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))
summary(mod)

sink()
```

Comparisons between means

```{r}
sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\n")

emm <- emmeans(mod, pairwise ~ condition)
emm.means <- summary(emm$emmeans, infer = T)
emm.contrasts <- summary(emm$contrasts, infer = T)
print(emm.means)
print(emm.contrasts)

sink()

emm_row <- function(df, col, value) df[df[[col]] == value, ]
macros <- macros %>%
  add_macros("expThreeEitherSocialDiff", emm_row_macros(emm_row(emm.contrasts, "contrast", "either refer - one social"))) %>%
  add_macros("expThreeReferSocialDiff", emm_row_macros(emm_row(emm.contrasts, "contrast", "one refer - one social")))
```

```{r}
# We want to check if "one social" is greater than  "one refer" and "either refer" together

sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\n")

emmeans_results <- emmeans(mod, ~ condition)
grouped_contrast <- contrast(emmeans_results,
                             method = list("one social - grouped refer" = c(0.5, 0.5, -1)),
                             adjust = "none")
summary(grouped_contrast, infer = T)

sink()
```

Broken down by tangram type

```{r}
sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\n")

emm.byshared <- emmeans(mod, pairwise ~ condition|shared)
emm.byshared.means <- summary(emm.byshared$emmeans, infer = T)
emm.byshared.contrasts <- summary(emm.byshared$contrasts, infer = T)
print(emm.byshared.means)
print(emm.byshared.contrasts)

sink()

# "unique" tangrams are the group-specific-label tangrams in the manuscript
emm_row_by <- function(df, col, value, shared_level) {
  df[df[[col]] == value & df$shared == shared_level, ]
}
macros <- macros %>%
  add_macros("expThreeGroupReferKnown", emm_row_macros(emm_row_by(emm.byshared.means, "condition", "one refer", "unique"))) %>%
  add_macros("expThreeGroupReferEither", emm_row_macros(emm_row_by(emm.byshared.means, "condition", "either refer", "unique"))) %>%
  add_macros("expThreeGroupEitherKnownDiff", emm_row_macros(emm_row_by(emm.byshared.contrasts, "contrast", "either refer - one refer", "unique"))) %>%
  add_macros("expThreeSharedReferKnown", emm_row_macros(emm_row_by(emm.byshared.means, "condition", "one refer", "shared"))) %>%
  add_macros("expThreeSharedReferEither", emm_row_macros(emm_row_by(emm.byshared.means, "condition", "either refer", "shared"))) %>%
  add_macros("expThreeSharedEitherKnownDiff", emm_row_macros(emm_row_by(emm.byshared.contrasts, "contrast", "either refer - one refer", "shared")))
```

```{r}
sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\n")

emmeans(mod, pairwise ~ condition*shared) %>%
  summary(infer = T)

sink()
```

## Sensitivity analysis

The key contrast (refer-known minus social-known, averaged over tangram type) under alternative random-effects specifications, from intercepts-only to maximal. The maximal specification is singular (boundary fit) and is shown only to demonstrate that the fixed-effects conclusions do not depend on the choice of structure.

```{r, eval = run_sensitivity}
sensitivity_specs <- list(
  "intercepts only" = response.earlier ~ condition * shared + (1 | subject_id) + (1 | item_id:tangram),
  "final (reported)" = response.earlier ~ condition * shared + (audience + shared | subject_id) + (1 | item_id:tangram),
  "maximal (singular)" = response.earlier ~ condition * shared + (condition * shared | subject_id) + (condition | item_id / tangram)
)

sensitivity <- map_dfr(names(sensitivity_specs), function(spec_name) {
  m <- glmer(sensitivity_specs[[spec_name]], data = main.trials, family = "binomial",
             control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))
  con <- summary(emmeans(m, pairwise ~ condition)$contrasts, infer = T) %>%
    as.data.frame() %>%
    filter(contrast == "one refer - one social")
  tibble(
    specification = spec_name,
    singular = isSingular(m),
    estimate = con$estimate, SE = con$SE,
    ci_lower = con$asymp.LCL, ci_upper = con$asymp.UCL,
    z = con$z.ratio, p = con$p.value
  )
})

write_csv(sensitivity, here("analysis/earlier-later/sensitivity.csv"))

sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\nSENSITIVITY ANALYSIS: refer-known minus social-known contrast (averaged\n")
cat("over tangram type) under alternative random-effects specifications (see\n")
cat("sensitivity.csv):\n\n")
print(as.data.frame(sensitivity))
sink()
```

## Baseline trials

For the "refer" goal, do participants successfully choose the seen label over the unseen label?

```{r}
sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\n")

# Following the preregistered simplification ladder: the maximal
# (1 | subject_id) + (1 | audience) + (1 | item_id/tangram) fit was singular
# because the tangram-set (item_id) and audience variances were both estimated
# at zero, so those components were removed. The intercept is unchanged to
# three decimals.
mod.same.refer <- glmer(seen ~ 1 + (1 | subject_id) + (1 | item_id:tangram), data = baseline.trials.refer, family = "binomial", control = glmerControl(optimizer = "bobyqa"))
summary(mod.same.refer)

sink()
```

For the "social" goal, do participants successfully choose the tangram that belongs to the audience group?

```{r}
sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\n")

mod.same.social <- glmer(group_label ~ 1 + (1 | subject_id) + (1 | item_id / tangram), data = baseline.trials.social, family = "binomial", control = glmerControl(optimizer = "bobyqa"))
summary(mod.same.social)

sink()
```

Convert intercepts from logit scale

```{r}
# Inverse logit function
inv_logit <- function(x) {
  exp(x) / (1 + exp(x))
}

sink(here("analysis/earlier-later/model_outputs.txt"), append = TRUE)
cat("\n\n")

cat("Baseline refer (inverse logit of intercept):\n")
print(inv_logit(fixef(mod.same.refer)))

cat("\n")

cat("Baseline social (inverse logit of intercept):\n")
print(inv_logit(fixef(mod.same.social)))

sink()
```

# Only look at participants who did above chance on the baselines

Filter for participants who did above chance on BOTH the social and refer baseline

```{r}
above_chance_same <- baseline.trials.refer %>%
  group_by(subject_id) %>%
  summarize(above_chance = mean(seen) > 0.5) %>%
  inner_join(baseline.trials.social %>%
               group_by(subject_id) %>%
               summarize(above_chance = mean(group_label) > 0.5), by = "subject_id") %>%
  filter(above_chance.x & above_chance.y) %>%
  select(subject_id)

selection_trials_filtered <- selection_trials %>%
  filter(subject_id %in% above_chance_same$subject_id)
```

How many participants?

```{r}
n_participants_filtered <- n_distinct(selection_trials_filtered$subject_id)
print(glue("we have {n_participants_filtered} participants who did above chance on both the social and refer baseline"))
```

Calculate means

```{r}
main.trials <- selection_trials_filtered %>%
  filter(type == "main") %>%
  mutate(response.earlier = ifelse(response.earlier == "earlier", 1, 0))

main.trials.means <- main.trials %>%
  group_by(condition, shared) %>%
  tidyboot_mean(response.earlier, na.rm = T)

# Per-participant means for the spaghetti lines; plain means (no bootstrap)
# because only the point estimates are used downstream.
main.trials.means.subj <-
  main.trials %>%
  group_by(subject_id, condition, shared) %>%
  summarize(empirical_stat = mean(response.earlier, na.rm = T), .groups = "drop")
```

Plot

```{r fig.width=8, fig.height=4}
ggplot(main.trials,
       aes(x = condition, y = response.earlier, color = shared)) +
  geom_point(
    data = main.trials.means,
    position = position_dodge(width = 0.3),
    aes(x = condition, y = empirical_stat, color = shared),
    size = 3,
    fill = "black",
    inherit.aes = FALSE
  ) +
  geom_errorbar(
    data = main.trials.means,
    position = position_dodge(width = 0.3),
    aes(
      x = condition,
      ymin = ci_lower,
      ymax = ci_upper,
      color = shared
    ),
    size = 1.5,
    width = 0.05,
    inherit.aes = FALSE
  ) +
  geom_line(
    data = main.trials.means.subj,
    aes(x = condition, y = empirical_stat, group = interaction(subject_id, shared), color = shared),
    position = position_dodge(width = 0.3),
    inherit.aes = FALSE,
    alpha = 0.15
  ) +
  geom_hline(yintercept = 0.5,
             color = "red",
             linetype = "dashed") +
  labs(x = "condition", y = "earlier or later label", title = "filtered participants") +
  scale_y_continuous(breaks = c(0, 1),
                     labels = c("later", "earlier")) +
  scale_color_discrete(name = "label", labels = c("shared", "group-specific"))
```



# Exit survey

What feedback did participants give?

```{r}
exit_survey$comments
```

# Export statistics as LaTeX macros

Write out every statistic quoted in the manuscript as a LaTeX macro, so the paper never hardcodes a computed value (`make manuscript-stats` copies these into the manuscript).

```{r}
write_macros(macros, here("analysis/earlier-later/macros.tex"))

# Export the final random-effects structure of each fitted model, straight
# from the fitted objects, for the manuscript's appendix table.
re_structures <- tibble::tibble(
  model = c("Label choice (main)", "Manipulation check: refer", "Manipulation check: social"),
  random_effects = vapply(list(mod, mod.same.refer, mod.same.social), re_terms, character(1))
)
write_csv(re_structures, here("analysis/earlier-later/re_structures.csv"))
```
