---
title: "paper_figs"
output: html_document
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(tidyboot)
library(glue)
library(showtext)
library(ggpubr)

theme_set(theme_classic(base_size = 24))
seaborn_colorblind <- c(
  "#0173b2", # blue
  "#de8f05", # orange
  "#029e73", # green
  "#d55e00", # red
  "#cc78bc", # purple
  "#ca9161", # brown
  "#fbafe4", # pink
  "#949494", # gray
  "#ece133", # yellow
  "#56b4e9"  # light blue
)

palette <- c(
  "Refer" = seaborn_colorblind[3],
  "Social" = seaborn_colorblind[2]
)
```

# Experiment 1


```{r}
data <- read_csv(here("data/free-response/selection_trials.csv")) %>%
  mutate(prop_naive = n_naive / (n_blue + n_naive),
         goal = if_else(goal == "refer", "Refer", "Social"))

data_filtered <- read_csv(here("data/free-response/selection_trials_filtered.csv")) %>%
  mutate(prop_naive = n_naive / (n_blue + n_naive),
         goal = if_else(goal == "refer", "Refer", "Social"))

demographics <- read_csv(here("data/free-response/exit_survey.csv"))
```

## Tangram selection

How often do people choose tangrams that belong to the in-group? 

```{r}
data <- data %>% 
  mutate(selected_out_group = if_else(selected_tangram_group == "red_specific", 1, 0))

data_filtered <- data_filtered %>% 
  mutate(selected_out_group = if_else(selected_tangram_group == "red_specific", 1, 0))
```

```{r}
tangram_selection_means <- data %>%
  group_by(prop_naive, goal) %>%
  tidyboot_mean(selected_out_group, na.rm = TRUE)

tangram_selection_means_filtered <- data_filtered %>%
  group_by(prop_naive, goal) %>%
  tidyboot_mean(selected_out_group, na.rm = TRUE)
```


Plot

```{r fig.width=6, fig.height=6}
ggplot(tangram_selection_means, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(position = position_dodge(width = 0.04), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.04), size = 2, width = 0) +
  scale_color_manual(values = palette) +
  labs(x = "Proportion naive", y = "P(selected out-group tangram)", color = "Goal") +
  theme(legend.position = "bottom")

ggsave(here("figures/outputs/tangram_selection.pdf"))
```

```{r fig.width=6, fig.height=6}
ggplot(tangram_selection_means_filtered, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(position = position_dodge(width = 0.04), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.04), size = 2, width = 0) +
  scale_color_manual(values = palette) +
  labs(x = "Proportion naive", y = "P(selected out-group tangram)", color = "Goal") +
  theme(legend.position = "bottom")

ggsave(here("figures/outputs/tangram_selection_filtered.pdf"))
```

##  Utterance length

Plot utterance length by proportion naive and goal

```{r}
utt_length_summary <- data %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(utt_length, na.rm = TRUE)

utt_length_summary_filtered <- data_filtered %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(utt_length, na.rm = TRUE)
```


```{r fig.width=6, fig.height=6}
ggplot(utt_length_summary, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(position = position_dodge(width = 0.04), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.04), size = 2, width = 0) +
  scale_color_manual(values = palette) +
  labs(x = "Proportion naive", y = "Utterance length", color = "Goal") +
  theme(legend.position = "bottom")

ggsave(here("figures/outputs/utterance_length.pdf"))
```
```{r fig.width=6, fig.height=6}
ggplot(utt_length_summary_filtered, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(position = position_dodge(width = 0.04), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.04), size = 2, width = 0) +
  scale_color_manual(values = palette) +
  labs(x = "Proportion naive", y = "Utterance length", color = "Goal") +
  theme(legend.position = "bottom")

ggsave(here("figures/outputs/utterance_length_filtered.pdf"))
```

## Cosine similarity to later utterance

Plot cosine similarity to later utterance by proportion naive and goal

```{r}
cos_diff_summary <- data %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(1 - sbert_cosine_later_blue, na.rm = TRUE)

cos_diff_summary_filtered <- data_filtered %>%
  group_by(goal, prop_naive) %>%
  tidyboot_mean(1 - sbert_cosine_later_blue, na.rm = TRUE)
```

```{r fig.width=6, fig.height=6}
ggplot(cos_diff_summary, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(position = position_dodge(width = 0.04), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.04), size = 2, width = 0) +
  scale_color_manual(values = palette) +
  labs(x = "Proportion naive", y = "Semantic distance to later", color = "Goal") +
  theme(legend.position = "bottom")

ggsave(here("figures/outputs/cosine_similarity.pdf"))
```

```{r fig.width=6, fig.height=6}
ggplot(cos_diff_summary_filtered, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(position = position_dodge(width = 0.04), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.04), size = 2, width = 0) +
  scale_color_manual(values = palette) +
  labs(x = "Proportion naive", y = "Semantic distance to later", color = "Goal") +
  theme(legend.position = "bottom")

ggsave(here("figures/outputs/cosine_similarity_filtered.pdf"))
```

# Experiment 2

```{r}
exp2.results <- read_csv(here("model/preds_exp2.csv")) %>%
  mutate(goal = ifelse(goal == "refer", "Refer", "Social"),
         audience = ifelse(audience == "either", "Either", "Known"),
         model = case_when(
           model == "base model" ~ "Base model",
           model == "social model" ~ "Social model",
           model == "data" ~ "Humans"
         )) %>%
  mutate(model = factor(model, levels = c("Humans", "Base model", "Social model")),
         audience = factor(audience, levels = c("Known", "Either")),
         goal = factor(goal, levels = c("Refer", "Social")))
```
```{r}
exp2.summary <- exp2.results %>% 
  group_by(audience, goal, model) %>% 
  tidyboot_mean(p_shared, na.rm = TRUE)
```

Plot

```{r fig.width=11, fig.height=3.8}
ggplot(exp2.summary, aes(x = audience, y = empirical_stat, color = goal)) +
  geom_hline(yintercept = 0.5, color = "gray", linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.4), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.4), size = 2, width = 0) +
  geom_line(aes(group = goal), position = position_dodge(width = 0.4), size = 2) +
  scale_color_manual(values = palette) +
  labs(x = "Audience group", y = "P(shared label)", color = "Goal", shape = "Model") +
  facet_wrap(~model)

ggsave(here("figures/outputs/exp2_results.pdf"))
```

# Experiment 3

```{r}
exp3.results <- read_csv(here("model/preds_exp3.csv")) %>%
  mutate(goal = ifelse(goal == "refer", "Refer", "Social"),
         audience = ifelse(audience == "either", "Either", "Known"),
         tangram_type = ifelse(tangram_type == "shared", "Shared", "Group-specific"),
         model = case_when(
           model == "base model" ~ "Base model",
           model == "social model" ~ "Social model",
           model == "data" ~ "Humans"
         )) %>%
  mutate(model = factor(model, levels = c("Humans", "Base model", "Social model")),
         audience = factor(audience, levels = c("Known", "Either")),
         tangram_type = factor(tangram_type, levels = c("Shared", "Group-specific")),
         goal = factor(goal, levels = c("Refer", "Social")))
```

```{r}
exp3.summary <- exp3.results %>% 
  group_by(audience, goal, tangram_type, model) %>% 
  tidyboot_mean(p_earlier, na.rm = TRUE)
```

calculate mean and SD for power analysis (later move this somewhere else)

```{r}
exp3.means <- exp3.results %>% 
  group_by(audience, goal, tangram_type, model) %>% 
  summarize(mean = mean(p_earlier, na.rm = TRUE),
            sd = sd(p_earlier, na.rm = TRUE))
```


```{r fig.width=11.5, fig.height=6}
ggplot(exp3.summary, aes(x = audience, y = empirical_stat, color = goal)) +
  geom_hline(yintercept = 0.5, color = "gray", linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.4), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.4), size = 2, width = 0) +
  geom_line(aes(group = goal), position = position_dodge(width = 0.4), size = 2) +
  scale_color_manual(values = palette) +
  labs(x = "Audience group", y = "P(earlier label)", color = "Goal", shape = "Model") +
  facet_grid(tangram_type ~ model)

ggsave(here("figures/outputs/exp3_results.pdf"))
```

# Experiment 4

```{r}
exp4.results <- read_csv(here("model/preds_exp4.csv")) %>%
  mutate(goal = ifelse(goal == "refer", "Refer", "Social"),
         model = case_when(
           model == "base model" ~ "Base model",
           model == "social model" ~ "Social model",
           model == "data" ~ "Humans"
         )) %>%
  mutate(model = factor(model, levels = c("Humans", "Base model", "Social model")),
         goal = factor(goal, levels = c("Refer", "Social")))
```
```{r}
exp4.summary <- exp4.results %>% 
  group_by(goal, model, prop_naive) %>% 
  tidyboot_mean(p_earlier, na.rm = TRUE)
```

Plot

```{r fig.width=12.5, fig.height=3.8}
ggplot(exp4.summary, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_hline(yintercept = 0.5, color = "gray", linetype = "dashed") +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(position = position_dodge(width = 0.05), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.05), size = 2, width = 0) +
  scale_color_manual(values = palette) +
  labs(x = "Proportion naive", y = "P(earlier label)", color = "Goal") +
  # legend on bottom 
  facet_wrap(~model)

ggsave(here("figures/outputs/exp4_results.pdf"))
```

## Experiment 4 transparency plots

```{r}
# Load transparency data
means_agg <- read_csv(here("data/transparency/means_agg.csv"))
means_by_label <- read_csv(here("data/transparency/means_by_label.csv"))

# Load exp4 selection trials
exp4.selection.trials <- read_csv(here("data/varied_audience/selection_trials.csv")) %>%
  rename(tangram_set = item_id) %>%
  mutate(response_coded = if_else(response.earlier == "earlier", 1, 0))
```

### Figure: Naive audience comprehension (earlier vs later)

```{r}
# Prepare data for bar plot
means_agg_plot <- means_agg %>%
  mutate(earlier = factor(earlier, levels = c("earlier", "later"),
                          labels = c("Earlier", "Later")))

# Get unique option pairs from exp4 main trials only
# In main trials, option1 is always "earlier" and option2 is always "later"
unique_option_pairs <- exp4.selection.trials %>%
  filter(type == "main") %>%
  select(tangram_set, option1.tangram, option1.label, option2.label) %>%
  distinct() %>%
  rename(target = option1.tangram,
         earlier_label = option1.label,
         later_label = option2.label)

# Join with means_by_label to get transparency estimates for earlier options
earlier_transparency <- means_by_label %>%
  filter(earlier == "earlier") %>%
  select(tangram_set, target, label, empirical_stat, ci_lower, ci_upper) %>%
  rename(earlier_label = label)

# Join with means_by_label to get transparency estimates for later options
later_transparency <- means_by_label %>%
  filter(earlier == "later") %>%
  select(tangram_set, target, label, empirical_stat, ci_lower, ci_upper) %>%
  rename(later_label = label)

# Join to get paired transparency data
itemwise_transparency <- unique_option_pairs %>%
  left_join(earlier_transparency, by = c("tangram_set", "target", "earlier_label")) %>%
  rename(earlier_stat = empirical_stat, earlier_ci_lower = ci_lower, earlier_ci_upper = ci_upper) %>%
  left_join(later_transparency, by = c("tangram_set", "target", "later_label")) %>%
  rename(later_stat = empirical_stat, later_ci_lower = ci_lower, later_ci_upper = ci_upper)

# Verify joins worked
cat("Number of unique option pairs:", nrow(unique_option_pairs), "\n")
cat("NAs in earlier_stat:", sum(is.na(itemwise_transparency$earlier_stat)), "\n")
cat("NAs in later_stat:", sum(is.na(itemwise_transparency$later_stat)), "\n")

# Create an item ID for each pair (for connecting lines)
itemwise_transparency <- itemwise_transparency %>%
  mutate(item_id = row_number())

# Pivot to long format for plotting
itemwise_long <- itemwise_transparency %>%
  select(item_id, earlier_stat, earlier_ci_lower, earlier_ci_upper, later_stat, later_ci_lower, later_ci_upper) %>%
  pivot_longer(
    cols = c(earlier_stat, later_stat),
    names_to = "type",
    values_to = "empirical_stat"
  ) %>%
  mutate(
    type = factor(if_else(type == "earlier_stat", "Earlier", "Later"), levels = c("Earlier", "Later")),
    ci_lower = if_else(type == "Earlier", earlier_ci_lower, later_ci_lower),
    ci_upper = if_else(type == "Earlier", earlier_ci_upper, later_ci_upper)
  ) %>%
  select(item_id, type, empirical_stat, ci_lower, ci_upper)

# Add jitter for x position
set.seed(42)
itemwise_long <- itemwise_long %>%
  mutate(x_jitter = as.numeric(type) + runif(n(), -0.08, 0.08))
```

```{r fig.width=5, fig.height=5}
ggplot() +
  geom_hline(yintercept = 1/6, color = "gray", linetype = "dashed") +
  # Bar plot of aggregate means
  geom_bar(data = means_agg_plot, aes(x = earlier, y = empirical_stat),
           stat = "identity", fill = "lightgray", width = 0.6) +
  geom_errorbar(data = means_agg_plot, aes(x = earlier, ymin = ci_lower, ymax = ci_upper),
                width = 0.2, linewidth = 1) +
  # Connecting lines for each item pair
  geom_line(data = itemwise_long, aes(x = x_jitter, y = empirical_stat, group = item_id),
            alpha = 0.3, color = "darkgray") +
  # Points with error bars for each item
  geom_point(data = itemwise_long, aes(x = x_jitter, y = empirical_stat),
             alpha = 0.2, size = 3, stroke = 0) +
  # geom_errorbar(data = itemwise_long, aes(x = x_jitter, ymin = ci_lower, ymax = ci_upper),
  #               width = 0.0, alpha = 0.3, size = 0.5) +
  labs(x = "Utterance type", y = "Proportion correct") +
  coord_cartesian(ylim = c(0, 1)) +
  theme(legend.position = "none")

ggsave(here("figures/outputs/transparency_earlier_later.pdf"))
```

### Figure: Predicted listener success by audience composition

```{r}
# Print a few rows to verify data
head(exp4.selection.trials)
head(means_by_label)
```

```{r}
# Join transparency estimates for option1, option2, and response
# First, prepare means_by_label for joining
transparency_lookup <- means_by_label %>%
  select(tangram_set, target, earlier, label, empirical_stat) %>%
  rename(transparency = empirical_stat)

# Verify a few rows of the lookup table
head(transparency_lookup)
```

```{r}
# Join option1 transparency
exp4.with.transparency <- exp4.selection.trials %>%
  left_join(
    transparency_lookup %>% rename(option1.transparency = transparency),
    by = c("tangram_set" = "tangram_set",
           "option1.tangram" = "target",
           "option1.earlier" = "earlier",
           "option1.label" = "label")
  )

# Verify join
cat("Option1 join check - rows with NA:", sum(is.na(exp4.with.transparency$option1.transparency)), "\n")
head(exp4.with.transparency %>% select(tangram_set, option1.tangram, option1.earlier, option1.label, option1.transparency))
```

```{r}
# Join option2 transparency
exp4.with.transparency <- exp4.with.transparency %>%
  left_join(
    transparency_lookup %>% rename(option2.transparency = transparency),
    by = c("tangram_set" = "tangram_set",
           "option2.tangram" = "target",
           "option2.earlier" = "earlier",
           "option2.label" = "label")
  )

# Verify join
cat("Option2 join check - rows with NA:", sum(is.na(exp4.with.transparency$option2.transparency)), "\n")
head(exp4.with.transparency %>% select(tangram_set, option2.tangram, option2.earlier, option2.label, option2.transparency))
```

```{r}
# Join response transparency
exp4.with.transparency <- exp4.with.transparency %>%
  left_join(
    transparency_lookup %>% rename(response.transparency = transparency),
    by = c("tangram_set" = "tangram_set",
           "response.tangram" = "target",
           "response.earlier" = "earlier",
           "response.label" = "label")
  )

# Verify join
cat("Response join check - rows with NA:", sum(is.na(exp4.with.transparency$response.transparency)), "\n")
head(exp4.with.transparency %>% select(tangram_set, response.tangram, response.earlier, response.label, response.transparency))
```

```{r}
# Calculate predicted listener success for each option and response
# Formula: (n_ingroup * 1 + n_outgroup * transparency) / (n_ingroup + n_outgroup)
exp4.with.transparency <- exp4.with.transparency %>%
  mutate(
    option1.pred.success = (n_ingroup * 1 + n_outgroup * option1.transparency) / (n_ingroup + n_outgroup),
    option2.pred.success = (n_ingroup * 1 + n_outgroup * option2.transparency) / (n_ingroup + n_outgroup),
    response.pred.success = (n_ingroup * 1 + n_outgroup * response.transparency) / (n_ingroup + n_outgroup),
    prop_naive = n_outgroup / (n_ingroup + n_outgroup)
  )

# Verify calculations
head(exp4.with.transparency %>% select(n_ingroup, n_outgroup, prop_naive, response.transparency, response.pred.success))
```

```{r}
# Group by prop_naive and goal, and tidyboot_mean response.pred.success
pred_success_summary <- exp4.with.transparency %>%
  filter(type == "main") %>%  # Only main trials
  mutate(goal = if_else(goal == "refer", "Refer", "Social")) %>%
  group_by(prop_naive, goal) %>%
  tidyboot_mean(response.pred.success, na.rm = TRUE)

head(pred_success_summary)
```

```{r fig.width=6, fig.height=6}
# Plot predicted listener success by audience composition and goal
ggplot(pred_success_summary, aes(x = prop_naive, y = empirical_stat, color = goal)) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(position = position_dodge(width = 0.02), size = 5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(width = 0.02), linewidth = 2, width = 0) +
  scale_color_manual(values = palette) +
  labs(x = "Proportion naive", y = "Predicted listener success", color = "Goal") +
  theme(legend.position = "bottom")

ggsave(here("figures/outputs/pred_listener_success.pdf"))
```

### Figure: Earlier label choice by transparency difference

```{r}
# For each trial, compute itemwise transparency difference (earlier option minus later option)
# Since option1 is always earlier and option2 is always later in the main trials,
# the transparency difference is option1.transparency - option2.transparency
exp4.with.transparency <- exp4.with.transparency %>%
  mutate(transparency_diff = option1.transparency - option2.transparency)

# Verify
head(exp4.with.transparency %>% select(option1.earlier, option2.earlier, option1.transparency, option2.transparency, transparency_diff))
```

```{r}
# Group by transparency difference and goal, and tidyboot_mean the response choice
# Round transparency_diff for binning
transparency_choice_summary <- exp4.with.transparency %>%
  filter(type == "main", prop_naive > 0) %>%
  mutate(goal = if_else(goal == "refer", "Refer", "Social"),
         transparency_diff_binned = round(transparency_diff, 2)) %>%
  group_by(transparency_diff_binned, goal) %>%
  tidyboot_mean(response_coded, na.rm = TRUE)

head(transparency_choice_summary)
```

```{r fig.width=6, fig.height=6}
# Plot: likelihood of choosing earlier label by transparency difference and goal
ggplot(transparency_choice_summary, aes(x = transparency_diff_binned, y = empirical_stat, color = goal)) +
  geom_smooth(method = "gam", linetype = "dashed", fill = "lightgray") +
  geom_point(size = 4, alpha = 0.7, stroke = 0) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0, linewidth = 1.3, alpha = 0.7) +
  geom_hline(yintercept = 0.5, color = "gray", linetype = "dashed") +
  scale_color_manual(values = palette) +
  labs(x = "Transparency diff (earlier > later)", y = "P(earlier label)", color = "Goal") +
  theme(legend.position = "bottom")

ggsave(here("figures/outputs/transparency_diff_choice.pdf"))
```






